```{r}
Sys.setlocale("LC_TIME", "English")
```


---
title: "PA1_template"
author: "Michal Svoboda"
date: "26 června 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting and processing the data 

activityData are downloaded from <href>https://d396qusza40orc.cloudfront.net/repactivityData%2FactivityData%2Factivity.zip</href>, unzipped
and loaded. The intervals are converted from integer representation to POSIXct representation.

```{r load,cache=TRUE}
#download.file("https://d396qusza40orc.cloudfront.net/repactivityData%2FactivityData%2Factivity.zip","activityactivityData")
#unzip("activityactivityData")
#commented out as the link stopped working

activityData<-read.csv("activity.csv")
activityData$time<-as.POSIXct(strptime(substr(as.character(activityData$interval+10000),2,5),format="%H%M"))

```

## Steps per day

Split the activityData into days and make a histogram. Also calculate mean and median. Note that missing
activityData are removed from the statistics.

```{r steps per day}
total.steps.per.day<-tapply(activityData$steps,activityData$date,sum,na.rm=FALSE)
hist(total.steps.per.day)
```


```{r,FALSE}
mean<-mean(total.steps.per.day,na.rm=TRUE)
median<-median(total.steps.per.day,na.rm=TRUE)
```
Mean is `r mean` and median is `r median`.

##steps.patternerage daily steps pattern

Count steps.patternerage steps in 5-minute intervals and make a plot of steps.patternerage daily pattern.
Also get the interval with highest number of steps.

```{r} 
library(dplyr)
steps.pattern<-summarize(group_by(activityData,time),steps=mean(steps,na.rm=TRUE))
plot(steps.pattern,type="l",main="Number of steps in 5-minute intervals")
#find at which time there were most steps recorded
steps.pattern$time[steps.pattern$steps==max(steps.pattern$steps)]
```

##Imputing missing values

Count the total number of missing values, choose imputing strategy (intervals with missing values 
will be replaced by a mean of this 5minute interval of all the valid days), show histogram of newly 
made dataset, calculate mean and median.
```{r imputing}
sum(is.na(activityData$steps))

#replace NAs by a typical 5minute inverval
impute<-function(interv){
    mean(activityData$steps[activityData$interval==interv],na.rm=TRUE)
}

activityData2<-activityData
activityData2$steps[is.na(activityData2$steps)]<-impute(activityData2$interval[is.na(activityData2$steps)])

total.steps.per.day2<-tapply(activityData2$steps,activityData2$date,sum,na.rm=FALSE)
hist(total.steps.per.day2)
mean2<-mean(total.steps.per.day2)
median2<-median(total.steps.per.day2)

```
Mean is `r mean2` and median is `r median2`. Not a big change from previous mean `r mean` and median `r median`.

##Weekend vs weekday

```{r}


#identify weekends or weekdays
activityData$weekday<-sapply(activityData$date,function(x){if(weekdays(as.Date(x))%in%c("Saturday","Sunday")){"Weekend"} else "Weekday"})

par(mfrow=c(2,1))
steps.pattern.weekend<-summarize(group_by(activityData[activityData$weekday=="Weekend",],time),steps=mean(steps,na.rm=TRUE))
steps.pattern.weekday<-summarize(group_by(activityData[activityData$weekday=="Weekday",],time),steps=mean(steps,na.rm=TRUE))

plot(steps.pattern.weekday,type="l",main="Weekday",ylim=c(0,250))
plot(steps.pattern.weekend,type="l",main="Weekend",ylim=c(0,250))

```

